---
title: "FNZ standard grids"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{FNZ standard grids}
  %\VignetteEncoding{UTF-8}
  %\VignetteEngine{knitr::rmarkdown}
editor_options: 
  chunk_output_type: console
---

```{r, include = FALSE}
knitr::opts_chunk$set(
  message = FALSE,
  collapse = TRUE,
  comment = "#>"
)
```

```{r echo=TRUE}
library(nzsf)
library(ggspatial)
library(stars)

theme_set(theme_bw() + theme(axis.title = element_blank()))
```

```{r EEZ_eg, fig.height=7, fig.width=7, fig.cap="The New Zealand EEZ (dashed black lines), a box bounding the EEZ (dashed red lines), a 50 x 50 km grid (blue lines), a 200 x 200 km grid (grey lines), and the origin (green point)."}
eez <- get_statistical_areas(area = "EEZ", proj = proj_nzsf()) 
bb_eez <- st_bbox(eez) %>% st_as_sfc()

grd200_eez <- get_standard_grid(cell_size = 200, 
                                bounding_box = st_bbox(eez), 
                                return_raster = FALSE)
grd050_eez <- get_standard_grid(cell_size = 50, 
                                bounding_box = st_bbox(eez), 
                                return_raster = FALSE)

# Plot and check with center point and bounding box
ggplot() +
  geom_sf(data = grd050_eez,  colour = "lightblue",  fill = NA, alpha = 0.15) +
  geom_sf(data = grd200_eez,  colour = "tomato",  fill = NA, alpha = 0.5) +
  plot_statistical_areas(area = "EEZ", colour = "black", fill = NA, linetype = "dashed") +
  geom_sf(data = bb_eez, colour = "red", fill = NA, linetype = "dashed") +
  plot_coast(resolution = "low", fill = "black", colour = "black", size = 0.3) +
  geom_point(aes(x = 0, y = -422600), colour = "green") + 
  annotation_scale(location = "tr", unit_category = "metric")
```

```{r CRA1_eg, fig.height=6, fig.width=7, fig.cap="The CRA 1 QMA (dashed black), a box bounding CRA 1 (dashed red), a 50 x 50 km grid (blue), and 200 x 200 km grid (grey). Note that there is no point of origin shown as it is outside of the CRA 1 QMA."}
cra1 <- get_statistical_areas(area = "CRA", proj = proj_nzsf()) %>% filter(QMA %in% "CRA1")
bb_cra1 <- st_bbox(cra1) %>% st_as_sfc()

grd200_cra1 <- get_standard_grid(cell_size = 200, 
                                 bounding_box = st_bbox(cra1), 
                                 return_raster = FALSE)
grd050_cra1 <- get_standard_grid(cell_size = 50, 
                                 bounding_box = st_bbox(cra1), 
                                 return_raster = FALSE)

ggplot() +
  geom_sf(data = grd050_cra1,  colour = "lightblue",  fill = NA, alpha = 0.5) +
  geom_sf(data = grd200_cra1,  colour = "tomato",  fill = NA, alpha = 0.5) +
  geom_sf(data = cra1, colour = "black", fill = NA, linetype = "dashed") +
  geom_sf(data = bb_cra1, colour = "red", fill = NA, linetype = "dashed") +
  plot_coast(resolution = "low", fill = "black", colour = "black", size = 0.3) +
  annotation_scale(location = "tr", unit_category = "metric") +
  plot_clip(x = grd200_cra1)
```

```{r overlap_eg, fig.height=7, fig.width=7, fig.cap="EEZ 200 x 200 km grid (blue), CRA 1 200 x 200 km grid (dashed red), origin (green point)."}
# Plot and check overlap of two grids
ggplot() +
  geom_sf(data = grd200_eez,  colour = "lightblue",  fill = NA, alpha = 0.5) +
  geom_sf(data = grd200_cra1,  colour = "blue",  fill = NA, alpha = 0.5, linetype = "dashed") +
  plot_coast(resolution = "low", fill = "black", colour = "black", size = 0.3) +
  geom_point(aes(x = 0, y = -422600), colour = "green") +
  annotation_scale(location = "tr", unit_category = "metric") +
  coord_sf(xlim = c(-4e+05, 2e+05), ylim = c(-422600, 777400))
```

```{r fine_eg, fig.height=3, fig.width=12, fig.cap="A 1 x 1 km grid (blue), origin (green point), and tip of Banks Peninsula to west. Note that the cell sizes agree well with the scale at top-right."}
bb1 <- st_bbox(eez)
bb1[1] <- -150000 # xmin
bb1[2] <- -422600 - 3000 # ymin (3 cells below the origin)
bb1[3] <- 3000 # xmax (3 cells to the right of the origin)
bb1[4] <- -400000 # ymax

grd001_eez <- get_standard_grid(cell_size = 1, 
                                bounding_box = bb1, 
                                return_raster = FALSE)

# Plot and check center point at fine scale
ggplot() +
  geom_sf(data = grd001_eez,  colour = "lightblue",  fill = NA, alpha = 0.15) +
  plot_coast(resolution = "low", fill = "black", colour = "black", size = 0.3) +
  geom_point(aes(x = 0, y = -422600), colour = "green") + 
  annotation_scale(location = "tr", unit_category = "metric") +
  plot_clip(x = grd001_eez)
```

```{r raster, fig.height=7, fig.width=7, fig.cap="EEZ 200 x 200 km grid as polygons (dotted red), EEZ 200 x 200 km grid as raster (blue), origin (green point)."}
# Get standard grid as a raster
r <- get_standard_grid(cell_size = 200, 
                       bounding_box = st_bbox(eez), 
                       return_raster = TRUE)

# Fill the grid with random values and plot it
r[] <- rnorm(n = ncell(r))
rstar <- st_as_stars(r)

ggplot() +
  geom_stars(data = rstar) +
  geom_sf(data = grd200_eez, fill = NA, colour = "red", linetype = "dotted") +
  plot_coast(resolution = "low", fill = "black", colour = "black", size = 0.3) +
  plot_statistical_areas(area = "EEZ", colour = "black", fill = NA) +
  geom_point(aes(x = 0, y = -422600), colour = "green")
```

In this next example I attempt to set up a grid of polygons and a raster that conforms to the document: A standardised approach for creating spatial grids for New Zealand marine environment and species data. I can almost get the polygons to work as I can specify the origin of {0, -422600} as an offset. However, this doesn't quite do what we want if the goal is to cover the New Zealand EEZ. I have no idea how to set up a raster with such an origin.

```{r echo=TRUE, fig.height=6, fig.width=9, message=FALSE}
# Working on A standardised approach for creating spatial grids for New Zealand marine environment and species data 
x <- raster(xmn = -1024479, xmx = 1091182, ymn = -1789355, ymx = 1545850)
# x <- raster(ncol = 36, nrow = 18, xmn = -1000, xmx = 1000, ymn = -100, ymx = 900)
projection(x) <- proj_nzsf()
origin(x) <- c(0, -422600)
origin(x)
res(x)
x

xp <- rasterToPoints(x) %>%
  data.frame()
xp$value <- rnorm(nrow(xp))

grd <- get_statistical_areas(area = "EEZ") %>%
  st_make_grid(cellsize = 500^2, offset = c(0, -422600), crs = proj_nzsf())

ggplot() +
  plot_statistical_areas(area = "EEZ") +
  geom_sf(data = grd) +
  geom_tile(data = xp, aes(x = x, y = y, fill = value), alpha = 0.5) +
  plot_coast(resolution = "low", fill = "black", colour = "black", size = 0.3) +
  geom_point(aes(x = 0, y = -422600), colour = "red") +
  annotation_scale(location = "tr", unit_category = "metric")
```
