---
title: "New Zealand Spatial Features"
output: 
  rmarkdown::html_vignette:
    toc: true
    toc_depth: 2
vignette: >
  %\VignetteIndexEntry{New Zealand Spatial Features}
  %\VignetteEncoding{UTF-8}
  %\VignetteEngine{knitr::rmarkdown}
editor_options: 
  chunk_output_type: console
---

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>"
)
```

The `nzsf` package relies heavily on the R packages `ggplot2`, `dplyr`, and `sf`. Maps can be built up in layers in the same way as `ggplot2`.

```{r echo=TRUE, message=FALSE}
# devtools::install_github(repo = "ropensci/rnaturalearthhires")
# library(rnaturalearthhires) # required for scale = "large" in ne_countries
library(rnaturalearth)
library(rnaturalearthdata)

library(nzsf)
library(tidyverse)
library(ggspatial)
library(viridis)
library(raster)
library(lwgeom)
library(patchwork)
library(stars)
library(ggnewscale)

theme_set(theme_bw() + theme(axis.title = element_blank()))
```

# CCAMLR

```{r echo=TRUE, fig.height=6, fig.width=9, message=FALSE}
# x <- gebco_CCAMLR %>%
#   st_as_stars() %>%
#   st_transform(crs = proj_ccamlr()) %>%
#   as.data.frame()
  # as("raster") %>%
  # as("SpatialPixelsDataFrame")
# x <- gebco_CCAMLR %>%
#   as("SpatialPixelsDataFrame") %>%
#   st_transform(crs = proj_ccamlr())

ggplot() +
  geom_ccamlr("mpa", fill = "tomato", colour = "tomato", alpha = 0.25) +
  geom_ccamlr("ssru") +
  geom_ccamlr("land", fill = "black") +
  geom_ccamlr("label") +
  annotation_north_arrow(location = "tr", which_north = "true", 
                         style = north_arrow_nautical, pad_y = unit(1, "cm")) +
  annotation_scale(location = "tr", unit_category = "metric") +
  coord_ccamlr()
```

# CCSBT

The `nzsf` package also includes functions for plotting CCSBT management areas. In the example below I simulate 100 points, generate a voronoi diagram around these points, then simulate values at 5000 points and sum the values of these points within the voronoi polygons:

```{r echo=TRUE, fig.width=11, message=FALSE}
CCSBT3994 <- CCSBT %>% 
  st_transform(proj_ccsbt()) %>%
  st_union(by_feature = TRUE)

# Simulate some points within the areas
pts1 <- st_sample(CCSBT3994, size = 100) %>% 
  st_sf()
pts2 <- st_sample(CCSBT3994, size = 5000) %>% 
  st_sf() %>% 
  mutate(z = rnorm(1:n()))

# Sum up points within voronoi polygons
vri <- pts1 %>%
  st_union() %>%
  st_voronoi(envelope = NULL) %>%
  st_collection_extract() %>%
  st_cast() %>%
  st_sf() %>%
  mutate(id = 1:n()) %>%
  st_join(pts2, join = st_contains, left = TRUE) %>%
  group_by(id) %>%
  summarise(z = sum(z)) %>%
  st_intersection(CCSBT3994)

ggplot() +
  geom_sf(data = vri, aes(fill = z), colour = NA) +
  scale_fill_viridis("Variable", alpha = 0.7, na.value = NA) +
  # plot_depth(proj = 3994) +
  geom_ccsbt("land", fill = "black", colour = "black") +
  geom_ccsbt("area", colour = "red") +
  geom_ccsbt("label", fill = "white", colour = "red") +
  coord_ccsbt()
```

```{r echo=TRUE, fig.height=6, fig.width=9, message=FALSE}
# register_google(key = key)
# 
# ggmap_bbox <- function(map, crs = 3857) {
#   if (!inherits(map, "ggmap")) stop("map must be a ggmap object")
#   # Extract the bounding box (in lat/lon) from the ggmap to a numeric vector, 
#   # and set the names to what sf::st_bbox expects:
#   map_bbox <- setNames(unlist(attr(map, "bb")), c("ymin", "xmin", "ymax", "xmax"))
# 
#   # Coonvert the bbox to an sf polygon, transform it to 3857, 
#   # and convert back to a bbox (convoluted, but it works)
#   bbox_3857 <- st_bbox(map_bbox, crs = 4326) %>%
#     st_as_sfc() %>%
#     st_transform(crs = crs) %>%
#     st_bbox()
# 
#   # Overwrite the bbox of the ggmap object with the transformed coordinates 
#   attr(map, "bb")$ll.lat <- bbox_3857["ymin"]
#   attr(map, "bb")$ll.lon <- bbox_3857["xmin"]
#   attr(map, "bb")$ur.lat <- bbox_3857["ymax"]
#   attr(map, "bb")$ur.lon <- bbox_3857["xmax"]
#   map
# }
# 
# 
# library(ggmap)
# get_map("Japan", zoom = 5, source = "stamen") %>% ggmap()
# 
# st_bbox(CCSBT %>% st_transform(crs = 4326))
# # gm <- get_map(location = "New Zealand", zoom = 2, source = "google", maptype = "satellite")
# gm <- get_map(location = c(lon = 0, lat = -30), zoom = 1, source = "google", maptype = "satellite")
# ggmap(gm)
# 
# 
# map <- ggmap_bbox(gm, crs = proj_ccsbt())
# # map <- ggmap_bbox(gm)
# # p <- ggmap(map) +
# p <- ggplot() +
#   annotation_map_tile(zoom = 7) +
#   geom_sf(data = CCSBT, fill = NA)
#   # coord_ccsbt()
# p
```

# SPRFMO

```{r echo=TRUE, fig.width=7, fig.height=7, message=FALSE}
SPRFMO3832 <- SPRFMO %>% 
  st_transform(crs = 3832)

world <- ne_countries(scale = "medium", returnclass = "sf") %>%
  st_transform(crs = 3832)

ggplot() +
  geom_sf(data = world, fill = "black", colour = "black") +
  geom_sf(data = SPRFMO3832, aes(fill = factor(OBJECTID))) +
  coord_sf() +
  plot_clip(SPRFMO3832) +
  theme_bw() +
  theme(legend.position = "none")
```

# SIOFA

```{r echo=TRUE, fig.width=7, fig.height=7, message=FALSE}
SIOFA3832 <- SIOFA %>%
  st_transform(crs = 3832)

world <- ne_countries(scale = "medium", returnclass = "sf") %>%
  st_transform(crs = 3832)

ggplot() +
  # geom_gebco(proj = 3832, downsample = 3) +
  # geom_stars(data = bathy, downsample = 3) +
  geom_sf(data = world, fill = "black", colour = "black") +
  new_scale("fill") +
  geom_sf(data = SIOFA3832, aes(fill = factor(SubArea), colour = factor(SubArea)), alpha = 0.5) +
  plot_clip(SIOFA3832) +
  theme_bw() +
  theme(legend.position = "none")
```

# New Zealand

```{r echo=TRUE, fig.height=6, fig.width=6, message=FALSE}
# Plot low resolution coastline
p1 <- ggplot() +
  plot_statistical_areas(area = "EEZ") +
  plot_coast(resolution = "1250k", fill = "black", colour = "black", size = 0.3) +
  plot_clip(x = "NZ") +
  annotation_north_arrow(location = "tl", which_north = "true", style = north_arrow_nautical) +
  annotation_scale(location = "br", unit_category = "metric")

# Get the coastline from the rnaturalearth package for comparison
nz <- ne_countries(scale = "medium", returnclass = "sf", country = "New Zealand") %>%
  st_transform(crs = proj_nzsf()) %>%
  st_crop(get_statistical_areas(area = "EEZ"))

p2 <- ggplot() +
  plot_statistical_areas(area = "EEZ") +
  geom_sf(data = nz, fill = "black", colour = "black") +
  annotation_north_arrow(location = "tl", which_north = "true", style = north_arrow_nautical) +
  annotation_scale(location = "br", unit_category = "metric")

p1 + p2
```

```{r echo=TRUE, fig.height=6, fig.width=6, message=FALSE}
# An example that aggregates spatial features
aa <- nz_general_statistical_areas %>%
  dplyr::select(Statistica) %>%
  st_transform(crs = proj_nzsf()) %>%
  st_union(by_feature = TRUE) %>%
  mutate(area = case_when(
    Statistica %in% c(401:412, "049", "050", "051", "052") ~ "a",
    Statistica %in% 601:625 ~ "b",
    TRUE ~ as.character("c")
  )) %>%
  group_by(area) %>%
  summarize(geometry = st_union(geometry))

ggplot() +
  geom_sf(data = aa, aes(fill = area)) +
  plot_qma(qma = "LIN", fill = "transparent") +
  # plot_statistical_areas(area = "stat area", fill = "transparent") +
  plot_coast(resolution = "med", fill = "forestgreen", colour = "black", size = 0.3) +
  plot_clip("NZ") +
  annotation_north_arrow(location = "tl", which_north = "true", style = north_arrow_nautical)
```

Layers such as New Zealand marine reserves, depth countours, and Quota Management Areas (QMAs) can be added easily with several of the `nzsf` helper functions including `plot_marine_reserves`, `plot_depth`, and `plot_qma`. Maps can be restricted (e.g. to the North Island only) using a bounding box generated using `st_bbox` from the `sf` package:

```{r echo=TRUE, fig.height=6, fig.width=6, message=FALSE}
# bbox <- get_coast() %>% 
bbox <- nzsf::nz_coastlines_and_islands_polygons_topo_1500k %>%
  st_transform(crs = proj_nzsf()) %>%
  filter(name %in% c("North Island or Te Ika-a-Māui")) %>%
  st_bbox()

ggplot() +
  plot_depth(colour = "lightblue") +
  plot_marine_reserves(fill = "red", colour = "red") +
  plot_qma(qma = "CRA", fill = NA) +
  plot_coast(resolution = "medium", fill = "grey", colour = NA, size = 0.3) +
  coord_sf(xlim = bbox[c(1, 3)], ylim = bbox[c(2, 4)]) +
  annotation_north_arrow(location = "tr", which_north = "true", style = north_arrow_nautical) +
  annotation_scale(location = "br", unit_category = "metric")
```

Adding labels can be done with:

```{r echo=TRUE, fig.height=6, fig.width=6, message=FALSE}
sf_jma <- get_qma("JMA")
sf_coast <- get_coast() %>% 
  st_combine() %>% 
  st_make_valid()
lab <- st_difference(sf_jma, sf_coast) %>% 
  st_point_on_surface()
# lab <- st_difference(sf_jma, sf_coast) %>% st_centroid()

ggplot() +
  plot_qma(qma = "JMA", fill = NA) +
  plot_statistical_areas(area = "JMA", fill = NA) +
  plot_coast(fill = "forestgreen", colour = NA, size = 0.3) +
  geom_sf_label(data = lab, aes(label = QMA)) +
  plot_clip(x = "NZ") +
  annotation_north_arrow(location = "tl", which_north = "true") +
  annotation_scale(location = "br", unit_category = "metric")
```

You can then add polygons, points, lines/arrows, and/or rasters to maps and change the map projection:

```{r echo=TRUE, fig.height=6, fig.width=6, message=FALSE}
proj <- "+proj=longlat +datum=WGS84 +no_defs"

data("Gisborne_TToR_Habitats")
Gisborne_TToR_Habitats <- Gisborne_TToR_Habitats %>% st_transform(crs = proj, check = TRUE)

data("Rocky_reef_National_NZ")
Rocky_reef_National_NZ <- Rocky_reef_National_NZ %>% st_transform(crs = proj, check = TRUE)

bbox <- get_marine_reserves() %>%
  st_transform(crs = proj, check = TRUE) %>%
  filter(Name == "Te Tapuwae o Rongokako Marine Reserve") %>%
  st_bbox()

ggplot() +
  geom_sf(data = Rocky_reef_National_NZ, fill = "lightgrey", colour = NA) +
  plot_depth(proj = proj, resolution = "med", size = 0.2, colour = "skyblue") +
  geom_sf(data = Gisborne_TToR_Habitats, aes(fill = Habitat), colour = NA) +
  scale_fill_viridis_d(alpha = 0.5) +
  plot_marine_reserves(proj = proj, fill = NA) +
  plot_coast(proj = proj, resolution = "med", fill = "black", colour = NA, size = 0.3) +
  # annotation_scale(location = "br", unit_category = "metric") +
  coord_sf(xlim = bbox[c(1, 3)], ylim = bbox[c(2, 4)]) +
  labs(title = "Te Tapuwae o Rongokako Marine Reserve")
```


```{r echo=TRUE, fig.height=6, fig.width=12, message=FALSE}
# stewart <- get_coast() %>%
stewart <- nzsf::nz_coastlines_and_islands_polygons_topo_1500k %>%
  filter(name == "Stewart Island/Rakiura") %>%
  st_transform(crs = proj_nzsf()) %>%
  st_buffer(dist = 4500)

# Simulate some points around Stewart Island
pts <- st_sample(stewart, size = 5000) %>% 
  st_sf() %>% 
  mutate(z = rnorm(1:n()))

p1 <- ggplot() +
  plot_depth(resolution = "med", size = 0.2, colour = "grey") +
  geom_sf(data = pts, aes(colour = z)) +
  plot_coast(resolution = "large", fill = "black", colour = NA, size = 0.3) +
  annotation_north_arrow(location = "tl", style = north_arrow_nautical) +
  # annotation_north_arrow(location = "tl", which_north = "true", style = north_arrow_nautical) +
  # annotation_scale(location = "br", unit_category = "metric") +
  plot_clip(x = stewart) +
  labs(colour = "Points", title = "Rakiura")

p2 <- ggplot() +
  plot_depth(resolution = "med", size = 0.2, colour = "grey") +
  plot_raster(data = pts, field = "z", fun = mean, nrow = 50, ncol = 50) +
  scale_fill_viridis("Raster", alpha = 0.8, option = "plasma") +
  plot_coast(resolution = "large", fill = "black", colour = NA, size = 0.3) +
  annotation_north_arrow(location = "tl", style = north_arrow_nautical) +
  # annotation_north_arrow(location = "tl", which_north = "true", style = north_arrow_nautical) +
  # annotation_scale(location = "br", unit_category = "metric") +
  plot_clip(x = stewart) +
  labs(title = "Rakiura")

p1 + p2
```

```{r echo=TRUE, fig.height=8, fig.width=8, message=FALSE}
eez <- get_statistical_areas("EEZ", proj = 4326) %>%
  st_shift_longitude()
r <- get_standard_grid(cell_size = 1/1000, bounding_box = st_bbox(eez), return_raster = FALSE, crs = 4326)

ggplot(data = eez) +
  geom_sf()

ggplot(data = r) +
  geom_sf(fill = "red", alpha = 0.1) +
  plot_coast(proj = 4326) +
  plot_clip(eez, proj = 4326) +
  theme_bw()

# library(ncdf4)
# library(ncdf.tools)
# library(ecmwfr)
# 
# wf_set_key(user = "youremail@gmail.com", key = "secretkey", service = "webapi")
# 
# request <- list(dataset      = "reanalysis-era5-single-levels-monthly-means",
#                 product_type = "reanalysis",
#                 variable     = "sea_surface_temperature",
#                 year         = 1979:2020,
#                 month        = 1:12,
#                 area         = "-30/160/-55/185",
#                 format       = "netcdf",
#                 target       = "era5-nz_sst_to_2020.nc")
# 
# ncfile <- wf_request(user = "41999", request = request, transfer = TRUE, path = "", verbose = FALSE)

# library(raster)
# # should add code to get this direct from web
# data(era5_nz_sst)
# prj <- "+proj=merc +lon_0=100 +lat_ts=-41 +x_0=0 +y_0=0 +datum=WGS84 +units=m +no_defs"
# 
# suppressWarnings({
#   sst1 <- era5_nz_sst %>% 
#     projectRaster(crs = prj) %>%
#     rasterToPoints()
# })
# 
# sst <- sst1 %>%
#   data.frame() %>%
#   pivot_longer(cols = c(-x, -y)) %>%
#   separate(col = name, into = c("year", "month", "day", "hr", "min", "sec"), sep = "\\.") %>%
#   mutate(year = gsub("X", "", year), value = value - 273.15) %>%
#   filter(year %in% c("1979", "1989", "2009", "2019"), month %in% c("01"))
# nz <- get_coast(proj = prj, resolution = "med")
# 
# ggplot() +
#   geom_tile(data = sst, aes(x = x, y = y, fill = value), alpha = 0.95) +
#   facet_wrap(year ~ month) +
#   geom_sf(data = nz, fill = "black", colour = NA, size = 0.3) +
#   coord_sf() +
#   scale_fill_viridis("SST (°C)", alpha = 0.8, option = "plasma")
```

```{r echo=TRUE, fig.height=6, fig.width=6, message=FALSE}
# f <- "../data-raw/mfe-average-seasurface-temperature-19932012-GTiff.zip"
# fz <- unzip(zipfile = f, list = TRUE)
# fz
# unzip(zipfile = f)

# mfe_average_sst <- raster::raster(x = "average-seasurface-temperature-19932012.tif", values = TRUE)# %>%
  # projectRaster(crs = proj_nzsf)
# names(mfe_average_sst) <- "layer"
# mfe_average_sst[mfe_average_sst[] < -10 | mfe_average_sst[] > 35] <- NA
# range(mfe_average_sst[], na.rm = TRUE)
# plot(mfe_average_sst)
# file.remove(fz$Name)
# use_data(mfe_average_sst, overwrite = TRUE)

  # st_transform(3994) %>%

# data(mfe_average_sst)

# rpts <- mfe_average_sst %>%
#   rasterToPoints() %>%
#   data.frame()

# ggplot() +
#   geom_raster(data = rpts, aes(x = x, y = y, fill = layer)) +
#   # plot_statistical_areas(area = "EEZ", fill = NA) +
#   # plot_coast(proj = projection(mfe_average_sst), resolution = "med", fill = "black", colour = NA, size = 0.3) +
#   coord_sf() +
#   scale_fill_viridis(alpha = 0.8, option = "magma") +
#   labs(fill = "SST (°C)")
```

```{r echo=TRUE, fig.height=6, fig.width=6, message=FALSE}
# data(gebco_depth_raster)
# 
# rpts <- gebco_depth_raster %>%
#   rasterToPoints() %>%
#   data.frame()
# 
# ggplot() +
#   geom_raster(data = rpts, aes(x = x, y = y, fill = layer)) +
#   plot_statistical_areas(area = "EEZ", fill = NA) +
#   plot_coast(resolution = "med", fill = "black", colour = NA, size = 0.3) +
#   coord_sf() +
#   scale_fill_viridis(alpha = 0.8, option = "magma") +
#   labs(fill = "Depth (m)")
```
