---
title: "New Zealand Spatial Features"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{nzsf}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>"
)
```

# Introduction

The `nzsf` package relies heavily on the R packages `tidyverse` and `sf`. Maps can be built up in layers in the same way as `ggplot2`. A basic map of the New Zealand coastline with a north arrow and a scale bar can be built using:

```{r echo=TRUE, fig.height=6, fig.width=6, message=FALSE}
library(nzsf)
library(ggspatial)
library(lwgeom)
library(viridis)

theme_set(theme_bw() + theme(axis.title = element_blank()))

ggplot() +
  plot_coast(resolution = "low", fill = "black", colour = "black", size = 0.3) +
  annotation_north_arrow(location = "tr", which_north = "true", style = north_arrow_nautical) +
  annotation_scale(location = "br", unit_category = "metric")
```

Layers such as depth, marine reserves, and Quota Management Areas (QMAs) can be added easily with several of the `nzsf` helper functions:

```{r echo=TRUE, fig.height=6, fig.width=6, message=FALSE}
proj <- "+proj=aea +lat_1=-30 +lat_2=-50 +lat=-40 +lon_0=175 +x_0=0 +y_0=0 +ellps=WGS84 +datum=WGS84 +units=m +no_defs"

bbox <- get_coast() %>% 
  st_transform(crs = proj, check = TRUE) %>% 
  filter(name %in% c("North Island or Te Ika-a-MÄui")) %>%
  st_bbox()

ggplot() +
  plot_depth(colour = "lightblue") +
  plot_marine_reserves(fill = "red", colour = "red") +
  plot_qma(qma = "CRA", fill = NA) +
  plot_coast(fill = "grey", colour = NA, size = 0.3) +
  annotation_north_arrow(location = "tr", which_north = "true", style = north_arrow_nautical) +
  annotation_scale(location = "br", unit_category = "metric") +
  coord_sf(xlim = bbox[c(1, 3)], ylim = bbox[c(2, 4)])
```

Adding labels can be a little tricky but can be done with:

```{r echo=TRUE, fig.height=6, fig.width=6, message=FALSE}
sf_jma <- get_qma("JMA")
sf_coast <- get_coast() %>% 
  st_transform(crs = proj, check = TRUE) %>% 
  st_combine() %>% 
  st_make_valid()
lab <- st_difference(sf_jma, sf_coast) %>% 
  st_point_on_surface()
# lab <- st_difference(sf_jma, sf_coast) %>% st_centroid()

ggplot() +
  plot_qma(qma = "JMA", fill = NA) +
  plot_statistical_areas(area = "JMA", fill = NA) +
  plot_coast(fill = "forestgreen", colour = NA, size = 0.3) +
  geom_sf_label(data = lab, aes(label = QMA)) +
  annotation_north_arrow(location = "tl", which_north = "true", style = north_arrow_fancy_orienteering) +
  annotation_scale(location = "br", unit_category = "metric")
```

You can then add polygons, points, lines, rasters, etc to maps:

```{r echo=TRUE, fig.height=6, fig.width=6, message=FALSE}
data("Gisborne_TToR_Habitats")
Gisborne_TToR_Habitats <- Gisborne_TToR_Habitats %>% st_transform(crs = proj, check = TRUE)

data("Rocky_reef_National_NZ")
Rocky_reef_National_NZ <- Rocky_reef_National_NZ %>% st_transform(crs = proj, check = TRUE)

bbox <- get_marine_reserves() %>%
  st_transform(crs = proj, check = TRUE) %>%
  filter(Name == "Te Tapuwae o Rongokako Marine Reserve") %>%
  st_buffer(dist = 4500) %>%
  st_bbox()

ggplot() +
  geom_sf(data = Rocky_reef_National_NZ, fill = "lightgrey", colour = NA) +
  plot_depth(resolution = "med", size = 0.2, colour = "skyblue") +
  geom_sf(data = Gisborne_TToR_Habitats, aes(fill = Habitat), colour = NA) +
  scale_fill_viridis_d(alpha = 0.5) +
  plot_marine_reserves(fill = NA) +
  plot_coast(resolution = "med", fill = "black", colour = NA, size = 0.3) +
  annotation_scale(location = "br", unit_category = "metric") +
  coord_sf(xlim = bbox[c(1, 3)], ylim = bbox[c(2, 4)])
```
